### 2019-01-04 methods.tex
grammar and clarity, mostly

@@

The JEFF-3.1.2 neutron library provided cross section generation 
\cite{oecd/nea_data_bank_jeff-3.1.2_2014}. 
The specific temperature was fixed 
for each material to correctly model the Doppler-broadening of resonance peaks 
when SERPENT generates the problem-dependent nuclear data library. The isotopic 
composition of each material at the initial state was described in detail in 
the MSBR conceptual design study \cite{robertson_conceptual_1971} and has been 
applied to the SERPENT model without any modification. Table~\ref{tab:msbr_tab} is 
a summary of the major \gls{MSBR} parameters used by this model 
\cite{robertson_conceptual_1971}.

@@

\subsubsection{Fuel material flows}
The $^{232}$Th in the fuel absorbs thermal neutrons and produces $^{233}$Pa 
which then decays into the fissile $^{233}$U. Furthermore, the \gls{MSBR} 
design requires online reprocessing to remove all poisons (e.g. $^{135}$Xe), 
noble metals, and gases (e.g. $^{75}$Se, $^{85}$Kr) every 20 seconds. 
Protactinium presents a challenge, since it has a large absorption cross 
section in the thermal energy spectrum. Moreover, $^{233}$Pa left in the core
 would produce $^{234}$Pa and $^{234}$U, which both are not useful as fuel, 
and smaller amount of $^{233}$Pa would decays into the fissile $^{233}$U.
Accordingly, $^{233}$Pa is continuously 
removed from the fuel salt into a protactinium decay tank to allow $^{233}$Pa 
to decay to $^{233}$U without negative impact on neutronics. The reactor 
reprocessing system must separate $^{233}$Pa from the molten-salt fuel over 3 
days, hold it while $^{233}$Pa decays into $^{233}$U, and return it back to the 
primary loop. This feature allows the reactor to avoid neutron losses to 
protactinium, lowers in-core fission product inventory, and increases the 
efficiency of $^{233}$U breeding. Table~\ref{tab:reprocessing_list} summarizes 
full list of nuclides and the ``cycle times''\footnote{ The \gls{MSBR} program defined a ``cycle time" as the amount of time required to remove 100\% of a target nuclide from a fuel 
salt \cite{robertson_conceptual_1971}.} used for modeling salt treatment and 
separations \cite{robertson_conceptual_1971}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{table}[ht!]
        \centering
        \caption{The effective cycle times for protactinium and fission 
        products removal (reproduced from \cite{robertson_conceptual_1971}).}
        \begin{tabularx}{\textwidth}{ x | s | x }
        \hline Processing group & \qquad\qquad\qquad Nuclides & Cycle time (at 
                full power) \\ \hline Rare earths & Y, La, Ce, Pr, Nd, Pm, Sm, 
                Gd & 50 days \\ \qquad & Eu & 500 days \\ Noble metals & Se, 
                Nb, Mo, Tc, Ru, Rh, Pd, Ag, Sb, Te & 20 sec \\
        Seminoble metals & Zr, Cd, In, Sn & 200 days \\
        Gases & Kr, Xe & 20 sec \\ Volatile fluorides & Br, I & 60 days \\
        Discard & Rb, Sr, Cs, Ba & 3435 days \\ 
        %Salt discard & Th, Li, Be, F & 3435 days \\ 
        Protactinium & $^{233}$Pa & 3 days \\ Higher 
                nuclides & $^{237}$Np, $^{242}$Pu & 16 years \\  \hline
        \end{tabularx}
        \label{tab:reprocessing_list}
\end{table}
The removal rates vary among nuclides in this reactor concept which dictate the 
necessary resolution of depletion calculations. If the depletion time intervals 
are very short, an enormous number of depletion steps are required to obtain 
the equilibrium composition. On the other hand, if the depletion  calculation 
time interval is too long, the impact of short-lived fission products is not 
captured. To compromise, the time interval for depletion calculations in this 
model was selected as 3 days\footnote{ Optimal depletion time step of 3 days for \gls{MSR} batch-wise depletion simulation was first described and concluded by Powers \emph{et al.} \cite{powers_new_2013}.} to correlate with the removal interval of 
$^{233}$Pa and $^{232}$Th was continuously added to maintain the initial mass 
fraction of $^{232}$Th.

@@

>>>>

$^{135}$Xe is a strong neutron poison, and some fraction of this gas 
is absorbed by graphite during \gls{MSBR} operation. ORNL calculations show 
that for unsealed commercial graphite with helium permeability 10$^{-5}$ 
cm$^2$/s the calculated poison fraction is less than 2\% 
\cite{robertson_conceptual_1971}.  This parameter can be improved by using 
experimental graphites or by applying sealing technology. The effect of the 
gradual poisoning of the core graphite with xenon is not treated here.

@@

The JEFF-3.1.2 neutron library provided cross section generation 
\cite{oecd/nea_data_bank_jeff-3.1.2_2014}. 
The specific temperature was fixed for each material and did not change during 
the reactor operation. The isotopic 
composition of each material at the initial state was described in detail in 
the MSBR conceptual design study \cite{robertson_conceptual_1971} and has been 
applied to the SERPENT model without any modification. Table~\ref{tab:msbr_tab} is 
a summary of the major \gls{MSBR} parameters used by this model 
\cite{robertson_conceptual_1971}.

@@

\subsubsection{Fuel material flows}
The $^{232}$Th in the fuel absorbs thermal neutrons and produces $^{233}$Pa 
which then decays into the fissile $^{233}$U. Furthermore, the \gls{MSBR} 
design requires online reprocessing to remove all poisons (e.g. $^{135}$Xe), 
noble metals, and gases (e.g. $^{75}$Se, $^{85}$Kr) every 20 seconds. 
Protactinium presents a challenge, since it has a large absorption cross 
section in the thermal energy spectrum. Moreover, $^{233}$Pa left in the core
 would produce $^{234}$Pa and $^{234}$U, neither of which are useful as fuel, 
 would produce a smaller amount of $^{233}$Pa which decays into the fissile $^{233}$U.
Accordingly, $^{233}$Pa is continuously 
removed from the fuel salt into a protactinium decay tank to allow $^{233}$Pa 
to decay to $^{233}$U without the corresponding negative neutronic impact. The reactor 
reprocessing system must separate $^{233}$Pa from the molten-salt fuel over 3 
days, hold it while $^{233}$Pa decays into $^{233}$U, and return it back to the 
primary loop. This feature allows the reactor to avoid neutron losses to 
protactinium, lowers in-core fission product inventory, and increases the 
efficiency of $^{233}$U breeding. Table~\ref{tab:reprocessing_list} summarizes 
full list of nuclides and the ``cycle times''\footnote{ The \gls{MSBR} program defined a ``cycle time" as the amount of time required to remove 100\% of a target nuclide from a fuel 
salt \cite{robertson_conceptual_1971}.} used for modeling salt treatment and 
separations \cite{robertson_conceptual_1971}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{table}[ht!]
        \centering
        \caption{The effective cycle times for protactinium and fission 
        products removal (reproduced from \cite{robertson_conceptual_1971}).}
        \begin{tabularx}{\textwidth}{ x | s | x }
        \hline Processing group & \qquad\qquad\qquad Nuclides & Cycle time (at 
                full power) \\ \hline Rare earths & Y, La, Ce, Pr, Nd, Pm, Sm, 
                Gd & 50 days \\ \qquad & Eu & 500 days \\ Noble metals & Se, 
                Nb, Mo, Tc, Ru, Rh, Pd, Ag, Sb, Te & 20 sec \\
        Seminoble metals & Zr, Cd, In, Sn & 200 days \\
        Gases & Kr, Xe & 20 sec \\ Volatile fluorides & Br, I & 60 days \\
        Discard & Rb, Sr, Cs, Ba & 3435 days \\ 
        %Salt discard & Th, Li, Be, F & 3435 days \\ 
        Protactinium & $^{233}$Pa & 3 days \\ Higher 
                nuclides & $^{237}$Np, $^{242}$Pu & 16 years \\  \hline
        \end{tabularx}
        \label{tab:reprocessing_list}
\end{table}
The removal rates vary among nuclides in this reactor concept which dictate the 
necessary resolution of depletion calculations. If the depletion time intervals 
are very short, an enormous number of depletion steps are required to obtain 
the equilibrium composition. On the other hand, if the depletion  calculation 
time interval is too long, the impact of short-lived fission products is not 
captured. To compromise, a 3 day time interval was selected for depletion 
calculations\footnote{ Optimal depletion time step of 3 days for \gls{MSR} 
batch-wise depletion simulation was first described and concluded by Powers 
\emph{et al.} \cite{powers_new_2013}.} to correlate with the removal interval 
of 
$^{233}$Pa and $^{232}$Th was continuously added to maintain the initial mass 
fraction of $^{232}$Th.

@@

The current version of SaltProc only allows 100\% separation efficiency for 
either specific elements or groups of elements (e.g. Processing Groups as described in 
Table~\ref{tab:reprocessing_list}) at the end of the specific cycle time. 
This approach works well for fast-removing elements (gases, noble metals, 
protactinium) which should be removed each depletion step. Unfortunately, 
for the elements with longer cycle times (i.e. rare earths should be removed 
every 50 days) this simplified approach leads to oscillatory behavior of all
major parameters. In future releases of SaltProc, this drawback will be eliminated 
by removing elements with longer cycle times using different method: only mass 
fraction (calculated separately for each reprocessing group) will be removed each
 depletion step (e.g. 3 days).

### 2019-01-04 results.tex
grammar and clarity, mostly

\section{Results}
The SaltProc online reprocessing simulation package is demonstrated in four 
applications: (1) analyzing  \gls{MSBR} neutronics and fuel cycle to find the 
equilibrium core composition and core depletion, (2) studying operational and 
safety parameters evolution during \gls{MSBR} operation, (3) demonstrating that 
in a single-fluid two-region \gls{MSBR} conceptual design the undermoderated 
outer core zone II works as a virtual ``blanket", reduces neutron leakage and 
improves breeding ratio due to neutron energy spectral shift, and (4) 
determining the effect of fission product removal on the core neutronics.

The neutron population per cycle and the number of active/inactive cycles were 
chosen to obtain balance between reasonable uncertainty for a transport problem 
($\leq$ 15 pcm\footnote{ 1 pcm = 10$^{-5}\Delta k_{eff}/k_{eff}$} for effective 
multiplication factor) and computational time. The \gls{MSBR} depletion and 
safety parameter computations were performed on 64 Blue Waters XK7 nodes (two 
AMD 6276 Interlagos CPU per node, 16 floating-point Bulldozer core units per 
node or 32 ``integer" cores per node, nominal clock speed is 2.45 GHz). The 
total computational time for calculating the equilibrium composition was 
approximately 9,900 node-hours (18 core-years.)

@@

Additionally, the presence of rubidium, strontium, cesium, and barium in the 
core are disadvantageous to reactor physics. In fact, SaltProc fully removes all these elements every 3435 days (not a small mass fraction every 3 days) 
which causes the multiplication factor to jump by approximately 450 
pcm, and limits using the batch approach for online reprocessing simulations. 
In future versions of SaltProc this drawback will be eliminated by removing 
elements with longer residence times (seminoble metals, volatile fluorides, Rb, Sr,
 Cs, Ba, Eu) using different approach. Only mass fraction (calculated 
 separately for each reprocessing group) will be removed every depletion step 
 (e.g. 3 days) instead of removing 100\% of element atoms after cycle time.
Overall, the effective multiplication factor gradually decreases from 1.075 to 
$\approx$1.02 at equilibrium after approximately 6 years of irradiation.

@@

\subsection{Neutron spectrum}
Figure~\ref{fig:spectrum} shows the normalized neutron flux spectrum for the 
full-core \gls{MSBR} model in the energy range from $10^{-8}$ to $10$ MeV. The 
neutron energy spectrum at equilibrium is harder than at startup due to 
plutonium and other strong absorbers accumulation in the core during reactor 
operation.  
\begin{figure}[ht!] % replace 't' with 'b'         to force it to 
  \centering
  \includegraphics[width=\textwidth]{spectrum.png} \caption{Neutron flux energy 
  spectrum normalized by unit lethargy and area under the curve is normalized to 1 for initial and equilibrium fuel salt 
  composition.}
  \label{fig:spectrum}
\end{figure}

@@

It is important to obtain the epithermal and thermal spectra to produce $^{233}$U from 
$^{232}$Th because the radiative capture cross section of thorium decreases monotonically 
from $10^{-10}$ MeV to $10^{-5}$ MeV. Hardening the spectrum tends to significantly 
increase resonance absorption in thorium and decrease absorptions in fissile and 
construction materials. 
\begin{figure}[ht!] % replace 't' with 'b' to force it to 
  \centering
  \includegraphics[width=\textwidth]{spectrum_zones.png} 
  \caption{Neutron flux energy spectrum in different core regions normalized by 
unit lethargy and area under the curve is normalized to 1 for the initial and equilibrium fuel salt composition.}
  \label{fig:spectrum_zones}
\end{figure}

@@

(table~\ref{tab:powgen_fraction}. 

@@

$^{233}$U production in the core. The thorium-233 produced after $^{232}$Th 
captured neutron then $\beta$-decays to 
$^{233}$Pa, which is the precursor for $^{233}$U production. Accordingly, this 

@@

Uncertainty for each temperature coefficient was obtain by propagating statistical 
 error of effective multiplication factor calculation 
provided by SERPENT2 and also appears in 
Table~\ref{tab:tcoef}. Other sources of uncertainty, such as cross section libraries 
uncertainty, error in salt and graphite density correlations, are not treated here.

@@

temperature feedback is an expansion of material that is heated. When the fuel 

@@

surfaces, was created based on this information. For the displacements calculation 
it was assumed that the interface between graphite reflector and vessel did not move,
 and the vessel temperature did not change. This is the most reasonable assumption for

@@

The fuel temperature coefficient (FTC) is negative for both initial and 
equilibrium fuel compositions due to thermal Doppler broadening of the resonance 
capture cross sections in the thorium. Small positive effect of fuel density on 
reactivity is increasing from +1.21 pcm/K for the reactor startup to +1.66 pcm/K for 
equilibrium fuel composition which has negative effect on FTC magnitude during the 
reactor operation. This is in good agreement with earlier 
research \cite{robertson_conceptual_1971,park_whole_2015}. The moderator 
temperature coefficient (MTC) is positive for the startup composition and decreases 
during reactor operation because of spectrum hardening with fuel depletion. 
Finally, the total temperature coefficient of reactivity is negative for both 
cases, but decreases during reactor operation due to spectral shift. In 
summary, even after 20 years of operation the total temperature coefficient of 
reactivity is relatively large and negative during reactor operation (comparing 
with conventional PWR which has temperature coefficient about -1.71 pcm/$^\circ$F 
$\approx$ -3.08 pcm/K \cite{forget_integral_2018}), despite positive MTC, and 
affords excellent reactor stability and control.

@@

Table~\ref{tab:six_factor} summarizes the six factors for both initial and 
equilibrium fuel salt composition. These factors and their statistical uncertainties
 have been calculated using SERPENT2 code for initial fuel salt composition (see 
Table~\ref{tab:msbr_tab}) and for equilibrium salt composition which was obtained 
with SaltProc. The non-leakage probability for both fast 
and thermal neutrons does not change during reactor operation because these 
values are not largely affected by the neutron spectrum shift. In contrast, 
neutron reproduction factor ($\eta$), resonance escape probability (p), and 
fast fission factor ($\epsilon$) are considerably different between startup and 
equilibrium. As indicated in Figure~\ref{fig:spectrum}, the neutron spectrum is 
softer at the beginning of reactor life. Neutron spectrum hardening causes the fast 
fission factor to increase through the core lifetime. The opposite is true for the 
resonance escape probability. Finally, the neutron reproduction factor 
decreases during reactor operation due to accumulation of fissile plutonium 
isotopes.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{table}[hb!]
  \centering
  \caption{Six factors for the full-core \gls{MSBR} model for initial and 
  equilibrium fuel composition.}
\begin{tabularx}{\textwidth}{ b | s | s } \hline
Factor  & Initial      & Equilibrium   \\ \hline
Neutron reproduction factor ($\eta$)     & $1.3960\pm.000052$     & 
        $1.3778\pm.00005$ \\ Thermal utilization factor (f)           & 
        $0.9670\pm.000011$     & $0.9706\pm.00001$ \\
Resonance escape probability (p)         & $0.6044\pm.000039$     & 
        $0.5761\pm.00004$ \\
Fast fission factor ($\epsilon$)         & $1.3421\pm.000040$     & 
        $1.3609\pm.00004$ \\
Fast non-leakage probability (P$_f$)     & $0.9999\pm.000004$     & 
        $0.9999\pm.000004$ \\
Thermal non-leakage probability (P$_t$)  & $0.9894\pm.000005$     & 
        $0.9912\pm.00005$ \\ \hline
\end{tabularx}
  \label{tab:six_factor}
\end{table}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Thorium refill rate}
In a \gls{MSBR} reprocessing scheme, the only external feed material flow  is 
$^{232}$Th. Figure~\ref{fig:th_refill} shows the $^{232}$Th feed rate 
calculated for 60 years of reactor operation. The $^{232}$Th feed rate 
fluctuates significantly as a result of the batch-wise nature of this online 
reprocessing approach. For example, the large spikes of up to 36 kg/day in a 
thorium consumption occurs every 3435 days. This is required due to batch-wise 
removal of strong absorbers (Rb, Sr, Cs, Ba). The corresponding effective 
multiplication factor increase (Figure~\ref{fig:keff}) and breeding 
intensification leads to additional $^{232}$Th consumption.  
\begin{figure}[ht!] % replace 't' with 'b' to force it to \centering
  \includegraphics[width=\textwidth]{Th_refill_rate.png} \caption{$^{232}$Th 
  feed rate over 60 years of \gls{MSBR} operation.}
  \label{fig:th_refill}
\end{figure}

The average thorium feed rate increases during the first 500 days of operation, 
and steadily decreases due to spectrum hardening and accumulation of 
absorbers in the core. As a result, the average $^{232}$Th feed rate over 60 
years of operation is about 2.40 kg/day. This thorium consumption rate is in 
good agreement with a recent online reprocessing study by \gls{ORNL} 
\cite{betzler_molten_2017}. It must be noted that for the reactor with only 
thorium feed, at near equilibrium state, the thorium consumption rate is 
determined by the reactor power, the energy released per fission, and 
neutron energy spectrum.

>>>>

\section{Results}
The SaltProc online reprocessing simulation package is demonstrated in four 
applications: (1) analyzing  \gls{MSBR} neutronics and fuel cycle to find the 
equilibrium core composition and core depletion, (2) studying operational and 
safety parameters evolution during \gls{MSBR} operation, (3) demonstrating that 
in a single-fluid two-region \gls{MSBR} conceptual design the undermoderated 
outer core zone II works as a virtual ``blanket'', reduces neutron leakage and 
improves breeding ratio due to neutron energy spectral shift, and (4) 
determining the effect of fission product removal on the core neutronics.

The neutron population per cycle and the number of active/inactive cycles were 
chosen to obtain balance between reasonable uncertainty for a transport problem 
($\leq$ 15 pcm\footnote{ 1 pcm = 10$^{-5}\Delta k_{eff}/k_{eff}$} for effective 
multiplication factor) and computational time. The \gls{MSBR} depletion and 
safety parameter computations were performed on 64 Blue Waters XK7 nodes (two 
AMD 6276 Interlagos CPU per node, 16 floating-point Bulldozer core units per 
node or 32 ``integer'' cores per node, nominal clock speed is 2.45 GHz). The 
total computational time for calculating the equilibrium composition was 
approximately 9,900 node-hours (18 core-years.)

@@

Additionally, the presence of rubidium, strontium, cesium, and barium in the 
core are disadvantageous to reactor physics. 
Overall, the effective multiplication factor gradually decreases from 1.075 to 
$\approx$1.02 at equilibrium after approximately 6 years of irradiation.

In fact, SaltProc fully removes 
all of these elements every 3435 days (not a small mass fraction every 3 days) 
which causes the multiplication factor to jump by approximately 450 
pcm, and limits using the batch approach for online reprocessing simulations. 
In future versions of SaltProc this drawback will be eliminated by removing 
elements with longer residence times (seminoble metals, volatile fluorides, Rb, Sr,
 Cs, Ba, Eu). In that approach, chemistry models will inform separation 
 efficiencies for each reprocessing group and removal will optionally be spread more 
 evenly accross the cycle time.

@@

\subsection{Neutron spectrum}
Figure~\ref{fig:spectrum} shows the normalized neutron flux spectrum for the 
full-core \gls{MSBR} model in the energy range from $10^{-8}$ to $10$ MeV. The 
neutron energy spectrum at equilibrium is harder than at startup due to 
plutonium and other strong absorbers accumulation in the core during reactor 
operation.  
\begin{figure}[ht!] % replace 't' with 'b'         to force it to 
  \centering
  \includegraphics[width=\textwidth]{spectrum.png} \caption{The neutron flux energy 
  spectrum is normalized by unit lethargy and the area under the curve is normalized to 1 for initial and equilibrium fuel salt 
  composition.}
  \label{fig:spectrum}
\end{figure}

@@

It is important to obtain the epithermal and thermal spectra to produce $^{233}$U from 
$^{232}$Th because the radiative capture cross section of thorium decreases monotonically 
from $10^{-10}$ MeV to $10^{-5}$ MeV. Hardening the spectrum tends to significantly 
increase resonance absorption in thorium and decrease absorptions in fissile and 
construction materials. 
\begin{figure}[ht!] % replace 't' with 'b' to force it to 
  \centering
  \includegraphics[width=\textwidth]{spectrum_zones.png} 
  \caption{The neutron flux energy spectrum in different core regions is normalized by 
unit lethargy and the area under the curve is normalized to 1 for the initial and equilibrium fuel salt composition.}
  \label{fig:spectrum_zones}
\end{figure}

@@

(table~\ref{tab:powgen_fraction}). 

@@

$^{233}$U production in the core. $^{232}$Th 
neutron capture produces $^{233}Th$ which then $\beta$-decays to 
$^{233}$Pa, the precursor for $^{233}$U production. Accordingly, this 

@@

By propagating the $k_{eff}$  statistical error provided by SERPENT2, 
uncertainty for each temperature coefficient was obtained and appears in 
Table~\ref{tab:tcoef}. Other sources of uncertainty are neglected, such as cross section 
measurement error and approximations inherent in the equations of state 
providing both the salt and graphite density dependence on temperature.

@@

temperature feedback is an expansion of heated material. When the fuel 

@@

surfaces, was created based on this information. For calculation of 
displacement, it was assumed that the interface between the graphite reflector and vessel did not move,
 and that the vessel temperature did not change. This is the most reasonable assumption for

@@

The fuel temperature coefficient (FTC) is negative for both initial and 
equilibrium fuel compositions due to thermal Doppler broadening of the resonance 
capture cross sections in the thorium. A small positive effect of fuel density on 
reactivity increases from $+1.21$ pcm/K at reactor startup to $+1.66$ pcm/K for 
equilibrium fuel composition which has negative effect on FTC magnitude during the 
reactor operation. This is in good agreement with earlier 
research \cite{robertson_conceptual_1971,park_whole_2015}. The moderator 
temperature coefficient (MTC) is positive for the startup composition and decreases 
during reactor operation because of spectrum hardening with fuel depletion. 
Finally, the total temperature coefficient of reactivity is negative for both 
cases, but decreases during reactor operation due to spectral shift. In 
summary, even after 20 years of operation the total temperature coefficient of 
reactivity is relatively large and negative during reactor operation (comparing 
with conventional PWR which has temperature coefficient about -1.71 pcm/$^\circ$F 
$\approx$ -3.08 pcm/K \cite{forget_integral_2018}), despite positive MTC, and 
affords excellent reactor stability and control.

@@

Table~\ref{tab:six_factor} summarizes the six factors for both initial and 
equilibrium fuel salt composition. Using SERPENT2 and SaltProc, tese factors and their statistical uncertainties
 have been calculated for both initial and equilibrium fuel salt composition (see 
Table~\ref{tab:msbr_tab}). The fast and thermal non-leakage probabilities 
remain constant despite the evolving neutron spectrum during operation. In 
contrast, the neutron reproduction factor ($\eta$), resonance escape 
probability ($p$), and fast fission factor ($\epsilon$) are considerably different between startup and 
equilibrium. As indicated in Figure~\ref{fig:spectrum}, the neutron spectrum is 
softer at the beginning of reactor life. Neutron spectrum hardening causes the fast 
fission factor to increase through the core lifetime. The opposite is true for the 
resonance escape probability. Finally, the neutron reproduction factor 
decreases during reactor operation due to accumulation of fissile plutonium 
isotopes.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{table}[hb!]
  \centering
  \caption{Six factors for the full-core \gls{MSBR} model for initial and 
  equilibrium fuel composition.}
\begin{tabularx}{\textwidth}{ b | s | s } \hline
Factor  & Initial      & Equilibrium   \\ \hline
Neutron reproduction factor ($\eta$)     & $1.3960\pm.000052$     & 
        $1.3778\pm.00005$ \\ Thermal utilization factor (f)           & 
        $0.9670\pm.000011$     & $0.9706\pm.00001$ \\
Resonance escape probability (p)         & $0.6044\pm.000039$     & 
        $0.5761\pm.00004$ \\
Fast fission factor ($\epsilon$)         & $1.3421\pm.000040$     & 
        $1.3609\pm.00004$ \\
Fast non-leakage probability (P$_f$)     & $0.9999\pm.000004$     & 
        $0.9999\pm.000004$ \\
Thermal non-leakage probability (P$_t$)  & $0.9894\pm.000005$     & 
        $0.9912\pm.00005$ \\ \hline
\end{tabularx}
  \label{tab:six_factor}
\end{table}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Thorium refill rate}
In a \gls{MSBR} reprocessing scheme, the only external feed material flow  is 
$^{232}$Th. Figure~\ref{fig:th_refill} shows the $^{232}$Th feed rate 
calculated for 60 years of reactor operation. The $^{232}$Th feed rate 
fluctuates significantly as a result of the batch-wise nature of this online 
reprocessing approach. For example, the large spikes of up to 36 kg/day in a 
thorium consumption occurs every 3435 days. This is required due to batch-wise 
removal of strong absorbers (Rb, Sr, Cs, Ba). The corresponding effective 
multiplication factor increase (Figure~\ref{fig:keff}) and breeding 
intensification leads to additional $^{232}$Th consumption.  
\begin{figure}[ht!] % replace 't' with 'b' to force it to \centering
  \includegraphics[width=\textwidth]{Th_refill_rate.png} \caption{$^{232}$Th 
  feed rate over 60 years of \gls{MSBR} operation.}
  \label{fig:th_refill}
\end{figure}

The average thorium feed rate increases during the first 500 days of operation, 
and steadily decreases due to spectrum hardening and accumulation of 
absorbers in the core. As a result, the average $^{232}$Th feed rate over 60 
years of operation is about 2.40 kg/day. This thorium consumption rate is in 
good agreement with a recent online reprocessing study by \gls{ORNL} 
\cite{betzler_molten_2017}. At equilibrium, the thorium feed rate is determined 
by the reactor power, the energy released per fission, and the neutron energy 
spectrum.
